B\'{e}zier $\bar{B}$ projection

### 2017-12-20 projection1120.tex
update

\begin{abstract}
  We demonstrate the use of B\'{e}zier projection to alleviate locking phenomena in structural mechanics applications of isogeometric analysis. We call this method B\'{e}zier $\bar{B}$ projection. To demonstrate the utility of the approach for both geometry and material locking phenomena we focus on transverse shear locking in Timoshenko beams and volumetric locking in nearly compressible linear elasticity although the approach can be applied generally to other types of locking phenemona as well. B\'{e}zier projection is a local projection technique with optimal approximation properties, which in many cases produces solutions that are comparable to global $L^2$ projection. In the context of $\bar{B}$ methods, the use of B\'ezier projection produces sparse stiffness matrices with only a slight increase in bandwidth when compared to standard displacement-based methods. Of particular importance is that the approach is applicable to any spline representation that can be written in B\'ezier form like NURBS, T-splines, LR-splines, etc. We discuss in detail how to integrate this approach into an existing finite element framework with minimal disruption through the use of B\'ezier extraction operators and a newly introduced dual \added{basis for} \Bezier \replaced{projection}{extraction} operator. We then demonstrate the behavior of the approach through several challenging benchmark problems.
\end{abstract}
\keywords{Isogeometric analysis, B\'ezier extraction, B\'ezier projection, $\bar{B}$-projection, locking}
\section{Introduction}
Isogeometric analysis (IGA), introduced by Hughes et al. \cite{hughes_isogeometric_2005}, adopts the spline basis, which underlies the CAD geometry, as the basis for analysis. Of particular importance is the positive impact of smoothness on numerical solutions, where, in many application domains, IGA outperforms classical finite elements~\cite{cottrell_isogeometric_2009,cottrell_studies_2007,cottrell2006isogeometric,hughes_duality_2008,bazilevs_isogeometric_2010,evans_n-widths_2009}. Initial investigations of IGA focused on non-uniform rational B-splines (NURBS) due to their dominance in commercial CAD packages. However, many advances are being made in analysis-suitable geometry representations that overcome the strict rectangular topological restrictions of NURBS. Examples include T-splines~\cite{bazilevs_isogeometric_2010,sederberg_t-splines_2003} and their analysis-suitable restriction~\cite{scott_local_2012, li_analysis-suitable_2013}, hierarchical B-splines~\cite{bornemann_subdivision-based_2013,scott_isogeometric_2014,schillinger_isogeometric_2012,evans_hierarchical_2015,forsey_hierarchical_1988}, and locally refined B-splines~\cite{dokken_polynomial_2013,johannessen_isogeometric_2014} among others.

@@

\subsection{Dual basis formulation of B\'{e}zier projection}
\label{sec:dual-funct-dual}
To integrate \Bezier projection into a standard finite element assembly algorithm, it is convenient to recast \Bezier projection in terms of a dual basis. A dual basis has the distinguishing property that
\begin{align}
  \int_{\Omega} \hat{N}_A N_B \, d\Omega = \delta_{AB}.
  \end{align}
Once a dual basis is defined it can be processed in much the same manner as standard basis functions are processed in a finite element code. A complete exposition on the subject of dual bases and the \Bezier projection framework can be found in~\cite{thomas_bezier_2015}. We first define the dual element extraction operator
\begin{align}
 \hat{\mathbf{D}}^e=\operatorname{diag}({\boldsymbol{\omega}^e})\mathbf{R}^e(\mathbf{G}^e)^{-1}
\end{align}
where $\mathbf{G}^e$ is the Gramian matrix of the Bernstein basis functions over the element and $\operatorname{diag}({\boldsymbol{\omega}^e})$ is a diagonal matrix that contains the B\'ezier projection weights computed by (\ref{eqn:Bezier_weight}). We can then define a dual basis function $\hat{N}_{A(e,a)}$ restricted to element $e$ as
\begin{align}
\hat{N}^e_{a} = \sum_{j}\hat{D}_{aj}^e B_{j}.
\label{eq:local-dual-basis}
\end{align}

@@

\added{The relationship of $\bar{B}$-formulation with mixed formulations has been studied and proved to be equivalent by Simo and Hughes \cite{hughes_variational_1986}. However, the symmetric \Bezier $\bar{B}$ projection lacks the connection to the mixed formulation. For this reason, instead of considering $\bar{B}$ formulation as a strain projection method, we regard it as a mixed formulation with the auxiliary variable ( the shear stress for the case of Timoshenko beam) being eliminated. Owing to the dual basis formulation of \Bezier projection operator, we use the \Bezier dual basis functions as the test functions of the auxiliary variable. Since the resulting stiffness matrix is not symmetric, this formulation is refered to as non-symmetric \Bezier $\bar{B}$ projection.}

>>>>

\begin{abstract}
  We demonstrate the use of B\'{e}zier projection to alleviate locking phenomena in structural mechanics applications of isogeometric analysis. \added{Based on two different 	
  interpretations of well-known $\bar{B}$ projection, we propose two formulations and name them (symmetric/non-symmetric) B\'{e}zier $\bar{B}$ projection.} To demonstrate the utility of the \replaced{\Bezier projection}{approach} for both geometry and material locking phenomena we focus on transverse shear locking in Timoshenko beams and volumetric locking in nearly compressible linear elasticity although the approach can be applied generally to other types of locking phenemona as well. B\'{e}zier projection is a local projection technique with optimal approximation properties, which in many cases produces solutions that are comparable to global $L^2$ projection. In the context of $\bar{B}$ methods, the use of B\'ezier projection produces sparse stiffness matrices with only a slight increase in bandwidth when compared to standard displacement-based methods. Of particular importance is that the approach is applicable to any spline representation that can be written in B\'ezier form like NURBS, T-splines, LR-splines, etc. We discuss in detail how to integrate this approach into an existing finite element framework with minimal disruption through the use of B\'ezier extraction operators and a newly introduced dual \added{basis for} \Bezier \replaced{projection}{extraction} operator. We then demonstrate the behavior of the \replaced{two proposed formulations}{approach} through several challenging benchmark problems.
\end{abstract}
\keywords{Isogeometric analysis, B\'ezier extraction, \added{\Bezier dual basis}, B\'ezier projection, $\bar{B}$-projection, locking}
\section{Introduction}
Isogeometric analysis (IGA), introduced by Hughes et al. \cite{hughes_isogeometric_2005}, adopts the spline basis, which underlies the CAD geometry, as the basis for analysis. Of particular importance is the positive impact of smoothness on numerical solutions, where, in many application domains, IGA outperforms classical finite elements~\cite{cottrell_isogeometric_2009,cottrell_studies_2007,cottrell2006isogeometric,hughes_duality_2008,bazilevs_isogeometric_2010,evans_n-widths_2009}. Initial investigations of IGA focused on non-uniform rational B-splines (NURBS) due to their dominance in commercial CAD packages. However, many advances are being made in analysis-suitable geometry representations that overcome the strict rectangular topological restrictions of NURBS. Examples include T-splines~\cite{bazilevs_isogeometric_2010,sederberg_t-splines_2003} and their analysis-suitable restriction~\cite{scott_local_2012, li_analysis-suitable_2013}, hierarchical B-splines~\cite{bornemann_subdivision-based_2013,scott_isogeometric_2014,schillinger_isogeometric_2012,evans_hierarchical_2015,forsey_hierarchical_1988}, and locally refined B-splines~\cite{dokken_polynomial_2013,johannessen_isogeometric_2014} among others.

@@

\subsection{Dual basis formulation of B\'{e}zier projection}
\label{sec:dual-funct-dual}
To integrate \Bezier projection into a standard finite element assembly algorithm, it is convenient to recast \Bezier projection in terms of a dual basis. \added{It plays a key role in developing the non-symmetric \Bezier $\bar{B}$ method.} A dual basis has the distinguishing property that
\begin{align}
  \int_{\Omega} \hat{N}_A N_B \, d\Omega = \delta_{AB}.
  \end{align}
Once a dual basis is defined it can be processed in much the same manner as standard basis functions are processed in a finite element code. A complete exposition on the subject of dual bases and the \Bezier projection framework can be found in~\cite{thomas_bezier_2015}. We first define the dual element extraction operator
\begin{align}
 \hat{\mathbf{D}}^e=\operatorname{diag}({\boldsymbol{\omega}^e})\mathbf{R}^e(\mathbf{G}^e)^{-1}
\end{align}
where $\mathbf{G}^e$ is the Gramian matrix of the Bernstein basis functions over the element and $\operatorname{diag}({\boldsymbol{\omega}^e})$ is a diagonal matrix that contains the B\'ezier projection weights computed by (\ref{eqn:Bezier_weight}). We can then define a dual basis function $\hat{N}_{A(e,a)}$ restricted to element $e$ as
\begin{align}
\hat{N}^e_{a} = \sum_{j}\hat{D}_{aj}^e B_{j}.
\label{eq:local-dual-basis}
\end{align}

@@

\added{The relationship of $\bar{B}$-formulation with mixed formulations has been studied and proved to be equivalent by Simo and Hughes \cite{hughes_variational_1986}. However, the symmetric \Bezier $\bar{B}$ projection lacks the connection to the mixed formulation. For this reason, instead of considering $\bar{B}$ formulation as a strain projection method, we regard it as a mixed formulation with the auxiliary variable (the shear stress for the case of Timoshenko beam) being eliminated. Taking the advantage of the dual basis formulation of \Bezier projection operator, we can incorporate \Bezier projection into the mixed formulation by using the \Bezier dual basis functions as the test functions of the auxiliary variable. Since the resulting stiffness matrix is not symmetric, this formulation is refered to as non-symmetric \Bezier $\bar{B}$ projection.}
